---
name: Mahi Srivastava, Pranav Jain
title: "R Notebook"
output: html_notebook
---

```{r}
library(DataComputing)
library(dplyr)
library(ggmap)
library(ggrepel)
library(tidyverse)

#Importing two different data sources.
airport_data <- read.csv("196400287_T_MASTER_CORD.csv")
routes_data <- read.csv("routes.csv")
```

### Inspecting Imported Data Sources
```{r}
str(airport_data)
```

```{r}
head(airport_data)
```

```{r}
str(routes_data)
```

```{r}
head(routes_data)
```

### General Data Wrangling 
```{r}
airport_data <-
  airport_data%>%
  rename(Airport_code = AIRPORT)
```

```{r}
airports <-
  airport_data%>%
  distinct(Airport_code, .keep_all = TRUE)
```

### Creating leaflet to show multiple airports on worldmap
```{r}
leaflet(airports)%>%
  addTiles()%>%
  addCircleMarkers(radius = 2,label = ~DISPLAY_AIRPORT_NAME)
```


### Join Operation for joining the two Data Sets by the column source.airport
```{r}
JointTable <-
  routes_data %>%
  left_join(airports %>% select(Airport_code,LATITUDE,LONGITUDE), by = c("source.airport" = "Airport_code"))
```

```{r}
JointTable <-
  JointTable%>%
  rename(source_latitude = LATITUDE,
         source_longitude = LONGITUDE)
```

### Join Operation for joining the two Data Sets by the column destination.apirport
```{r}
JointTable <-
  JointTable %>%
  left_join(airports %>% select(Airport_code,LATITUDE,LONGITUDE), by = c("destination.apirport" = "Airport_code"))
```

```{r}
JointTable <-
  JointTable%>%
  rename(destination_latitude = LATITUDE,
         destination_longitude = LONGITUDE)
```

### Inspecting the JointTable
```{r}
str(JointTable)
```

###Removing the unnecessary columns from JointTable
```{r}
JointTable <-
  JointTable %>%
  select(airline, source.airport, destination.apirport, equipment, source_latitude, source_longitude, destination_latitude, destination_longitude)
```


###Creating a user - defined function
```{r}
earth.dist <- function (long1, lat1, long2, lat2)
{
rad <- pi/180
a1 <- lat1 * rad
a2 <- long1 * rad
b1 <- lat2 * rad
b2 <- long2 * rad
dlon <- b2 - a2
dlat <- b1 - a1
a <- (sin(dlat/2))^2 + cos(a1) * cos(b1) * (sin(dlon/2))^2
c <- 2 * atan2(sqrt(a), sqrt(1 - a))
R <- 6378.145
d <- R * c
return(d)
}
```

### Finding the distance between two coordinates
```{r}
JointTable<-
  JointTable%>%
  mutate(distance = earth.dist(source_longitude,source_latitude,destination_longitude,destination_latitude))
```

### Grouping the distance in short, edium or long with the use of Control Flow (if-else statement)
```{r}
JointTable <-
  JointTable%>%
  mutate(Distance_class = ifelse(distance <= 3000, "short", ifelse(distance >3000 & distance <= 9000, "Medium", "Long")))
```

### Making a table for all the long routes with the use of reduction/transformation function
```{r}
LongTable <-
  JointTable%>%
  filter(Distance_class == "Long")%>%
  group_by(equipment)%>%
  summarise(count = n())%>%
  arrange(desc(count))%>%
  head(10)
```

### Plotting a Graph to show which aircraft model (equipment) is most common for long flights 
```{r}
ggplot(data=LongTable,aes(x=equipment,y=count ))+geom_bar(stat='identity',position='stack', width=.9)
```

The above graph tells us that 777 is the most used aircraft model(equipment) for long flights. 

### Making a new table which has two variables that show the aircraft model(equipment) and the number of routes it is used for (count)
```{r}
routes_777 <- 
  JointTable%>%
  filter(Distance_class == "Long", equipment == "777")
```

### Plotting a map to show routes of long flights with the help of multiple different geoms and mutiple aesthetics
```{r}
worldmap <- borders("world", colour = "#FFFFFF", fill = "#efede1") 
  ggplot() + worldmap + 
    geom_point(data = routes_777, aes(x = source_longitude, y = source_latitude), col = "#970027") + 
    geom_curve(data = routes_777, aes(x = source_longitude, y = source_latitude, xend = destination_longitude, yend = destination_latitude), 
               col = "#b29e7d", size = .4) + 
    theme_void()
```


### Making a table for all the medium routes with the use of reduction/transformation function
```{r}
MediumTable <-
  JointTable%>%
  filter(Distance_class == "Medium")%>%
  group_by(equipment)%>%
  summarise(count = n())%>%
  arrange(desc(count))%>%
  head(10)
```

### Plotting a Graph to show which aircraft model (equipment) is most common for medium flights
```{r}
ggplot(data=MediumTable,aes(x=equipment,y=count ))+geom_bar(stat='identity',position='stack', width=.9)
```
The above graph tells us that 320 is the most used aircraft model(equipment) for medium flights.

### Making a new table which has two variables that show the aircraft model(equipment) and the number of routes it is used for(count)
```{r}
routes_320 <- 
  JointTable%>%
  filter(Distance_class == "Medium", equipment == "320")
```

### Plotting a map to show routes of medium flights with the help of multiple different geoms and mutiple aesthetics
```{r}
worldmap <- borders("world", colour = "#FFFFFF", fill = "#efede1") 
  ggplot() + worldmap + 
    geom_point(data = routes_320, aes(x = source_longitude, y = source_latitude), col = "#970027") + 
    geom_curve(data = routes_320, aes(x = source_longitude, y = source_latitude, xend = destination_longitude, yend = destination_latitude), 
               col = "#b29e7d", size = .4) + 
    theme_void()
```

### Making a table for all the short routes with the use of reduction/transformation function
```{r}
ShortTable <-
  JointTable%>%
  filter(Distance_class == "short")%>%
  group_by(equipment)%>%
  summarise(count = n())%>%
  arrange(desc(count))%>%
  head(10)
```

### Plotting a Graph to show which aircraft model (equipment) is most common for short flights
```{r}
ggplot(data=ShortTable,aes(x=equipment,y=count ))+geom_bar(stat='identity',position='stack', width=.9)
```

The above graph tells us that 320 is the most used aircraft model(equipment) for short flights.

### Making a new table which has two variables that show the aircraft model(equipment) and the number of routes it is used for(count)
```{r}
routes_320_short <- 
  JointTable%>%
  filter(Distance_class == "short", equipment == "320")
```

### Plotting a map to show routes of short flights with the help of multiple different geoms and mutiple aesthetics
```{r}
worldmap <- borders("world", colour = "#FFFFFF", fill = "#efede1") 
  ggplot() + worldmap + 
    geom_point(data = routes_320_short, aes(x = source_longitude, y = source_latitude), col = "#970027") + 
    geom_curve(data = routes_320_short, aes(x = source_longitude, y = source_latitude, xend = destination_longitude, yend = destination_latitude), 
               col = "#b29e7d", size = .4) + 
    theme_void()
```

